---
title: Bayesian Estimation from Gaussian Distribution ($\mu$,$\lambda$)
author: Akihiko Mori
date: '2022-05-17'
slug: []
categories: []
tags: 
- Bayes Statistics
---

## Purpose
I will implement Bayesian estimation of the parameter $\mu$ and $\lambda$ of a Gaussian (Normal) distribution using R. 
For the prior distribution, specify Gaussian and Gamma distributions and seek the posterior distribution. 
Implemented from the creation of data, it can be copied and pasted.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data
5. Posterior Distribution

## Library
```{r}
library(ggplot2)
set.seed(1)
```

## True Distribution
```{r}
mu <- 170
lambda <- 1/100
x <- seq(100 ,200, 1)
prob <- dnorm(x, mu, 1/sqrt(lambda))

ggplot()+
  geom_line(aes(x=x, y=prob))+
  labs(x="x", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 20))+
  ggeasy::easy_center_title()
```


## Prior Distribution (Gamma)
```{r}
a_pre <- 2
b_pre <- 1
x <- seq(0 ,0.1, 0.001)
lambda_prob_pre <- dgamma(x, a_pre, b_pre)
lambda_sample_pre <- rgamma(1, a_pre, b_pre)

plot_lambda <- ggplot()+
  geom_line(aes(x=x, y=lambda_prob_pre))+
  labs(x="λ", y="prob", title="Gamma Distribution")+
  theme_classic(base_family = "")+
  theme(text = element_text(size = 20))+
  ggeasy::easy_center_title()
plot_lambda
```
## Prior Distribution (Gaussian)
```{r}
mu_pre <- 150
beta_pre <- 0.0001
x <- seq(100 ,200, 1)
mu_prob_pre <- dnorm(x, mu_pre, 1/sqrt(beta_pre*lambda_sample_pre))

plot_mu <- ggplot()+
  geom_line(aes(x=x, y=mu_prob_pre))+
  labs(x="μ", y="prob", title="Normal (Prior)")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 20))+
  ggeasy::easy_center_title()
plot_mu
```

## Data
```{r}
data <- rnorm(30, mu, 1/sqrt(lambda))
N <- length(data)
```

## Bayesian Update
```{r}
beta_pos <- N + beta_pre
mu_pos <- (sum(data)+beta_pre*mu_pre)/beta_pos
a_pos <- N/2 + a_pre
b_pos <- (sum(data^2)+beta_pre*mu_pre^2-beta_pos*mu_pos^2)/2 + b_pre
lambda_sample_pos <- rgamma(1, a_pos, b_pos)
```



## Post Distribution (Gamma)
```{r}
x <- seq(0 ,0.1, 0.001)
lambda_prob_pos <- dgamma(x, a_pos, b_pos)

plot_lambda + geom_line(aes(x=x, y=lambda_prob_pos), col="blue")
```

## Post Distribution (Gaussian)
```{r}
x <- seq(100 ,200, 1)
mu_prob_pos <- dnorm(x, mu_pos, 1/sqrt(beta_pos*lambda_sample_pos))

plot_mu + geom_line(aes(x=x, y=mu_prob_pos), col="blue")
```