---
title: Bayesian Estimation from Multivariate Normal Distribution ($\mu$)
author: Akihiko Mori
date: '2022-05-17'
slug: []
categories: []
tags: 
- Bayes Statistics
---

## Purpose
Bayesian estimation of the parameter $\mu$ from multivariate Gaussian distribution (multivariate normal distribution) will be implemented using R. 
The prior distribution is assumed as a multidimensional Gaussian distribution and 
seek the posterior distribution. 
Implemented from the generation of data, the R code can be copied and pasted.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r}
library(ggplot2)
library(MCMCpack)
set.seed(1)
```

## True Distribution
```{r}
mu <- c(170, 60)
lambda <- solve(matrix(c(100, 80, 80, 100), ncol = 2))
samples <- mvrnorm(100, mu = mu, Sigma = solve(lambda))

ggplot()+
  geom_point(aes(x=samples[,1], y=samples[,2]))+
  labs(x="x", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 20))+
  lims(x=c(140,200), y=c(30,90))+
  ggeasy::easy_center_title()
```


## Prior Distribution
```{r}
mu_pre <- c(150, 50)
lambda_pre <- solve(matrix(c(1000, 100, 100, 1000), ncol = 2))
mu_samples_pre <- mvrnorm(100, mu = mu_pre, Sigma = solve(lambda_pre))

plot <- ggplot()+
  geom_point(aes(x=mu_samples_pre[,1], y=mu_samples_pre[,2]))+
  labs(x=expression(mu[1]), y=expression(mu[2]), title="Gaussian Distribution")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 20))+
  lims(x=c(40,250), y=c(-50,160))
plot
```

## Data Generation
```{r}
data <- mvrnorm(30, mu = mu, Sigma = solve(lambda))
N <- nrow(data)
```

## Post Distribution
```{r}
lambda_pos <- N*lambda + lambda_pre
mu_pos <- solve(lambda_pos) %*% (lambda %*% apply(data, 2, sum) + lambda_pre %*% mu_pre)
mu_samples_pos <- mvrnorm(100, mu = mu_pos, Sigma = solve(lambda_pos))

plot + geom_point(aes(x=mu_samples_pos[,1], y=mu_samples_pos[,2]), col="blue")
```