---
title: Bayeian Estimation with Multivariate Normal Distribution (mu)
author: Akihiko Mori
date: '2022-06-22'
slug: []
categories: []
tags:
  - Bayesian Statistics (R)
---

## Purpose
Bayesian estimation of the parameter μ of Multivariate Gaussian (Normal) distribution will be implemented using R. 
For the prior distribution, Multivariate Gaussian distributions will be specified, and the posterior distribution will be obtained. 
This will implement this from data generation, so it can be reproduced by copy and paste.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r message=FALSE}
library(ggplot2)
library(MCMCpack)
set.seed(1)
```

## True Distribution
```{r, message=FALSE}
mu <- c(170, 60)
lambda <- solve(matrix(c(100, 80, 80, 100), ncol = 2))
samples <- mvrnorm(100, mu = mu, Sigma = solve(lambda))

ggplot()+
  geom_point(aes(x=samples[,1], y=samples[,2]))+
  labs(x="x", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))+
  lims(x=c(140,200), y=c(30,90))
```


## Prior Distribution
```{r, message=FALSE}
mu_pre <- c(150, 50)
lambda_pre <- solve(matrix(c(1000, 100, 100, 1000), ncol = 2))
mu_samples_pre <- mvrnorm(100, mu = mu_pre, Sigma = solve(lambda_pre))

plot <- ggplot()+
  geom_point(aes(x=mu_samples_pre[,1], y=mu_samples_pre[,2]))+
  labs(x=expression(mu[1]), y=expression(mu[2]), title="Gaussian Distributions")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))+
  lims(x=c(40,250), y=c(-50,160))
plot
```


## Data Generation
```{r}
data <- mvrnorm(30, mu = mu, Sigma = solve(lambda))
N <- nrow(data)
```

## Posterior Distribution
### Gaussian Distribution
```{r}
lambda_pos <- N*lambda + lambda_pre
mu_pos <- solve(lambda_pos) %*% (lambda %*% apply(data, 2, sum) + lambda_pre %*% mu_pre)
mu_samples_pos <- mvrnorm(100, mu = mu_pos, Sigma = solve(lambda_pos))

plot + geom_point(aes(x=mu_samples_pos[,1], y=mu_samples_pos[,2]), col="blue")
```
