---
title: Normal Linear Model with BRMS
author: Akihiko Mori
date: '2022-05-28'
slug: []
categories: []
tags: 
- Bayesian Modelling (R and Stan)
---

## Purpose
Bayesian estimation of the GLM, a normal linear model, will be implemented using R and Stan. This code is implemented with the library of BRMS.
The data will be implemented from the data preparation and can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. BRMS
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(brms) # BRMS
library(patchwork) 
library(margins)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Normal Linear Model | $\mu_n=\beta_0+\beta_1x_{n1}+\cdots +\beta_{M}x_{nM}$ <br />$p(y_n | \beta,\sigma) \sim Normal(y_n|\mu_n,\sigma^2)$<br /> |
| This Model          | $\mu_n=\beta_0+\beta_1Age_n+\beta_2Qual_n$ <br />$p(Income_n | \beta,\sigma) \sim Normal(Income_n | \mu_n,\sigma^2)$<br />   |
| Simulation          |$\mu_n=300+10Age_n+200Qual_n$ <br />$Income_n \sim Normal(Income_n | \mu_n,100^2)$<br /> |

## Data Generation
```{r}
qual <- rbinom(100, 1, 0.3)
age <- rnorm(100, 50, 20) %>% round()
income <- rnorm(100, 300 + 10*age + 200*qual, 100) %>% round()
data <- data.frame(qual, age, income) %>% filter(age>18, age<65)
data %>% head()

plot <- ggplot() +
  geom_point(aes(x=data$age, y=data$income, color=factor(data$qual))) + 
  theme_classic(base_family = "") +
  theme(text=element_text(size=20))+
  labs(x="Age", y="Icome", title="Scatter plot") +
  scale_color_manual("Qualification",values=c("red","blue"))+
  ggeasy::easy_center_title()
plot
```

## BRMS
```{r}
mcmc_result <- brm(
  data = data,
  formula = income ~ age + qual,
  family = gaussian(link = identity),
  seed = 1,
  iter = 1000, warmup = 200, chains = 4, thin=1
)
```


## Analysis Results
### R code
```{r}
# Result
print(mcmc_result)

## Convergence Check
theme_set(theme_classic(base_size = 10, base_family = ""))
plot(mcmc_result)

## mu result
plot(conditional_effects(mcmc_result, 
                         effects="age:qual"), 
     points=TRUE)%>%wrap_plots()+plot_annotation(title="μ's Estimation")

## Prediction distribution
plot(conditional_effects(mcmc_result, 
                         effects="age:qual", 
                         method="predict"),
     points=TRUE) %>%wrap_plots() +
  plot_annotation(title="prediction distribution")
```