---
title: Bayeian Estimation  with Categorical Distribution
author: Akihiko Mori
date: '2022-06-21'
slug: []
categories: []
tags:
  - Bayesian Statistics (R)
---

## Purpose
Bayesian estimation of the parameter p of the Categorical distribution will be implemented using R. 
For the prior distribution, a beta distribution will be specified, and the posterior distribution will be obtained. 
We will implement this from data generation, so it can be reproduced by copy and paste.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r message=FALSE}
library(ggplot2)
library(MCMCpack)
set.seed(1)
```

## True Distribution
```{r}
x <- c(1, 2, 3)
p <- c(0.1, 0.6, 0.3)
prob <- p

ggplot() + 
  geom_bar(aes(x, prob), stat ="identity") +
  labs(x="x", y="prob", title="Categorical Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))
```


## Prior Distribution
```{r}
alpha_pre <- c(1, 1, 1)
p_samples_pre <- rdirichlet(100, alpha_pre)

plot <- ggplot()+
  geom_point(aes(x=p_samples_pre[,1], y=p_samples_pre[,2]))+
  labs(x="1", y="2", title="Dirichlet Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))+
  lims(x=0:1, y=0:1)
plot
```


## Data Generation
```{r}
data <- which(rmultinom(300, 1, c(0.1, 0.6, 0.3))==1, arr.ind = TRUE)[,1]
N <- length(data)
```

## Posterior Distribution

```{r}
alpha_pos <- as.vector(table(data)) + alpha_pre
p_samples_pos <- rdirichlet(100, alpha_pos)

plot + geom_point(aes(x=p_samples_pos[,1], y=p_samples_pos[,2]), col="blue")
```