---
title: Bayeian Estimation with Multivariate Gaussian Distribution (μ,λ)
author: Akihiko Mori
date: '2022-06-22'
slug: []
categories: []
tags:
  - Bayesian Statistics (R)
---

## Purpose
Bayesian estimation of the parameter μ,λ of Multivariate Gaussian (Normal) distribution will be implemented using R. 
For the prior distribution, Multivariate Gaussian and Wishart distribution will be specified, and the posterior distribution will be obtained. 
This will implement this from data generation, so it can be reproduced by copy and paste.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r message=FALSE}
library(ggplot2)
library(ellipse)
library(MCMCpack)
set.seed(1)
```

## True Distribution
```{r, message=FALSE}
mu <- c(170, 60)
lambda <- solve(matrix(c(100, 80, 80, 100), ncol = 2))
samples <- mvrnorm(100, mu = mu, Sigma = solve(lambda))

ell <- ellipse(centre=mu, solve(lambda))
ggplot()+
  geom_point(aes(x=samples[,1], y=samples[,2]))+
  labs(x=expression(x[1]), y=expression(x[2]), title="Multivariate Gaussian Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 20))+
  lims(x=c(140,200), y=c(30,90))+
  geom_polygon(data=as.data.frame(ell), aes(x,y), fill=NA,colour="black")
```


## Prior Distribution
### Wishart distribution (λ)
```{r, message=FALSE}
n_pre <- 3
w_pre <- diag(2)*0.01
lambda_sample_pre <- rwish(n_pre, w_pre)

plot_lambda <- ggplot() +
  theme_classic(base_family = "HiraKakuPro-W3")+
  theme(text = element_text(size = 20), legend.position=c(0.3,0.7)) +
  labs(x=expression(x[1]), y=expression(x[2]), title="Wishart distribution") 
for(i in 1:100){
  lambda_samples_pre <- rwish(n_pre, w_pre)
  ell <-  ellipse(centre=mu, solve(lambda_samples_pre))
  plot_lambda <- plot_lambda + geom_polygon(data=as.data.frame(ell),aes(x,y), fill=NA,colour="black")
}
plot_lambda
```
### Multivariate Gaussian Distribution (μ)
```{r}
mu_pre <- c(150, 50)
beta_pre <- 0.01
mu_samples_pre <- mvrnorm(100, mu = mu_pre, Sigma = solve(beta_pre*lambda_sample_pre))

plot_mu <- ggplot()+
  geom_point(aes(x=mu_samples_pre[,1], y=mu_samples_pre[,2]))+
  labs(x=expression(mu[1]), y=expression(mu[2]), title="Multivariate Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))
plot_mu
```


## Data Generation
```{r}
data <- mvrnorm(100, mu = mu, Sigma = solve(lambda))
N <- nrow(data)
```

## Posterior Distribution
### Parameters update
```{r}
beta_pos <- N + beta_pre
mu_pos <- (apply(data,2,sum)+beta_pre*mu_pre)/beta_pos
n_pos <- N + n_pre
S <- 0
for(i in 1:N){
  S <- S + data[i,] %*% t(data[i,])
}
w_pos <- solve(S + beta_pre*mu_pre%*%t(mu_pre) - beta_pos*mu_pos%*%t(mu_pos) + solve(w_pre))
lambda_sample_pos <- rwish(n_pos, w_pos)
```

### Wishart distribution (λ)
```{r}
for(i in 1:100){
  lambda_samples_pos <- rwish(n_pos, w_pos)
  ell <-  ellipse(centre=mu, solve(lambda_samples_pos))
  plot_lambda <- plot_lambda + geom_polygon(data=as.data.frame(ell),aes(x,y), fill=NA, colour="blue")
}
plot_lambda
```


### Gaussian Distribution
```{r}
mu_samples_pos <- mvrnorm(100, mu = mu_pos, Sigma = solve(beta_pos*lambda_sample_pos))

plot_mu + geom_point(aes(x=mu_samples_pos[,1], y=mu_samples_pos[,2]), col="blue")
```