---
title: 'ENSC650 lecture 01: Vector and Matrix Algebra'
author: Akihiko Mori
date: '2021-09-15'
slug: []
categories: []
tags:
  - R
---

## Vector
A vector is a quantity that has both **magnitude** and **direction**.
In general, a vector is a column vector.  

```{r}
a <- c(1,2,3)
b <- c(2,1,3)
```
The order is mattered, $a \neq b$.

### Transpose of Vector
Transposing a vector means turning a column (row) vector into a row (column) vector.

$$\left[\begin{array}
{r}
1  \\
2  \\
3 
\end{array}\right]^T
=
\left[\begin{array}
{rrr}
1 & 2 & 3
\end{array}\right]
$$

```{r}
aT <- t(a)
```

## Matrices
A rectangular array of numbers/ functions.
$M\times N$ matrix $\mathbf{A}$ is a table with $m$ rows and $n$ columns; or  size of $\mathbf{A}$ is $M\times N$. Each element is $a_{ij}, i\in M, j\in N$.
$$\mathbf{A_{MN}} = \left[\begin{array}
{rrrr}
a_{11} & a_{12} & \dots & a_{1n} \\
a_{21} & a_{22} &  & \vdots  \\
\vdots &  & \ddots & \vdots \\
a_{m1} & \dots & \dots  & a_{mn} 
\end{array}\right]
$$

```{r}
# matrix - column by column is default
A <- matrix(c(1,2,3,
              4,5,6,
              7,8,9),3,3)
# row by row matrix assign
B <- matrix(c(1,2,3,
              4,5,6,
              7,8,9),3,3, byrow = TRUE)
A # column by column
B # row by row
```

### Transpose of Matrix
Transposing matrix is interchanging rows and columns. $90^\circ$ rotation.
$$a_{ij} = a_{ji}$$

e.g.
$$\left[\begin{array}
{rrr}
1 & 3 & 8 \\
4 & 7 & 2 \\
6 & 5 & 0
\end{array}\right]^T
= \left[\begin{array}
{rrr}
1 & 4 & 6 \\
3 & 7 & 5 \\
8 & 2 & 0
\end{array}\right]
$$

```{r}
AT <- t(A)
```

### Special Matrices
#### Diagonal Mtrix
A square matrix with 0 on all off–diagonal entries and elements $d_1,d_2,\dots,d_n$
on the diagonal a diagonal matrix,  $diag{d_1, d_2,\dots , d_n}$

$$\left[\begin{array}
{rrr}
1 & 0 & 0\\
0 & 2 & 0\\
0 & 0 & 3
\end{array}\right]$$

```{r}
diag(c(1,2,3))
```

– A diagonal matrix with $1$s on the diagonal is the identity matrix
and is denoted $\mathbf{I}$. The identity matrix satisfies that $\mathbf{IA} = \mathbf{AI} = \mathbf{A}$.

$$\left[\begin{array}
{rrr}
1 & 0 & 0\\
0 & 1 & 0\\
0 & 0 & 1
\end{array}\right]$$

```{r}
diag(1,3)
```

#### Symmetric Matrix
A matrix is symmetric if $$\mathbf{A}^T=\mathbf{A}$$

e.g.  
$$ \left[\begin{array}
{rr}
1 & 2 \\
2 & 3 
\end{array}\right]^T
= \left[\begin{array}
{rr}
1 & 2 \\
2 & 3 
\end{array}\right]
$$

#### Skew-Symmetric Matrix
A matrix is skew-symmetric if $$\mathbf{A}^T=-\mathbf{A}$$

e.g.
$$\left[\begin{array}
{rrr}
0 & 2 & -45 \\
-2 & 0 & -4 \\
45 & 4 & 0
\end{array}\right]^T
= \left[\begin{array}
{rrr}
0 & -2 & 45 \\
2 & 0 & 4 \\
-45 & -4 & 0
\end{array}\right]
$$

#### Orthogonal Matrix
A matrix is orthogonal if $$\mathbf{A}^T = \mathbf{A}^{-1}$$

e.g.
$$ \left[\begin{array}
{rr}
1 & 0 \\
0 & 1 
\end{array}\right]^T
= \left[\begin{array}
{rr}
1 & 0 \\
0 & 1 
\end{array}\right],
 \left[\begin{array}
{rr}
1 & 0 \\
0 & 1 
\end{array}\right]^{-1}
= \left[\begin{array}
{rr}
1 & 0 \\
0 & 1 
\end{array}\right]
$$

$$ \left[\begin{array}
{rr}
cos\theta & -sin\theta \\
sin\theta & cos\theta 
\end{array}\right]
$$

$$ \left[\begin{array}
{rr}
1 & 0 \\
0 & -1 
\end{array}\right]
$$

$$ \left[\begin{array}
{rrrr}
0 & 0  & 0 & 1 \\
0 & 0  & 1 & 0 \\
1 & 0  & 0 & 0 \\
0 & 1  & 0 & 0 
\end{array}\right]
$$

### Scalar Multiplication

For a number $\alpha$ and a matrix $\mathbf{A}$, the product $\alpha\mathbf{A}$ is the matrix obtained by multiplying each element in $\mathbf{A}$ by $\alpha$.

e.g.
$$5\times\left[\begin{array}
{rr}
1 & 3 \\
4 & 7 
\end{array}\right] =
\left[\begin{array}
{rr}
5 & 15 \\
20 & 35 
\end{array}\right]
$$

```{r}
A <- matrix(c(1,3,4,7),2,2,byrow = T)
a <- 5
a*A
```

### Addition
Let $\mathbf{A}$ and $\mathbf{B}$ be $m\times n$ matrices. The sum $$\mathbf{A+B}$$ is the $m\times n$ matrix obtained by adding $\mathbf{A}$ and $\mathbf{B}$ elementwise.
Only matrices with the same dimensions can be added.

e.g.
$$\left[\begin{array}
{rr}
1 & 3 \\
4 & 7 
\end{array}\right] +
\left[\begin{array}
{rr}
6 & 2 \\
3 & 1 
\end{array}\right] 
=
\left[\begin{array}
{rr}
7 & 5 \\
7 & 8 
\end{array}\right]
$$

```{r}
A <- matrix(c(1,3,4,7),2,2,T)
B <- matrix(c(6,2,3,1),2,2,T)
A+B
```

### Multiplication
Let $\mathbf{A}$ be a $m\times l$ matrix and $\mathbf{B}$ a $l\times n$ matrix, 
i.e.$B=[b_1 :b_2 :···:b_t]$. The product $\mathbf{AB}$ is the $m \times n$ matrix given by:
$$\mathbf{AB}=\mathbf{A}\left[b_1 :b_2 :\dots:b_t\right]=\left[\mathbf{A}b_1 :\mathbf{A}b_2 :\dots:\mathbf{A}b_t\right]$$
$$\mathbf{C}_{mn}=\sum_{l=1}^L\mathbf{A}_{ml}\mathbf{B}_{ln}$$
in general $$\mathbf{AB} \neq \mathbf{BA}$$

e.g.
$$\left[\begin{array}
{rr}
4 & 3 \\
7 & 2 \\
9 & 0
\end{array}\right]
\left[\begin{array}
{rr}
2 & 5 \\
1 & 6 
\end{array}\right] 
=
\left[\begin{array}
{rr}
4\cdot 2 + 3\cdot 1  & 4\cdot 5+3\cdot6 \\
7\cdot 2 + 2\cdot 1  & 7\cdot 5+2\cdot6 \\
9\cdot 2 + 0\cdot 1  & 9\cdot 5+0\cdot6 
\end{array}\right] = 
\left[\begin{array}
{rr}
11 & 38 \\
16 & 47 \\
18 & 45
\end{array}\right]
$$

```{r}
A <- matrix(c(4,3,7,2,9,0),3,2,T)
B <- matrix(c(2,5,1,6),2,2,T)
# Matrix Multiplication
A%*%B
```

### Inverse of Matrices - Division
The inverse of an $n \times n$ matrix $\mathbf{A}$ is the matrix $\mathbf{B}$ (which is also $n \times n$) which multiplied with $\mathbf{A}$ generates the identity matrix $\mathbf{I}$:
$$\mathbf{AB} = \mathbf{BA} = \mathbf{I}$$
$\mathbf{B}$ is an inverse of the matrix $\mathbf{A}$, called $\mathbf{A^{-1}}$.


```{r}
solve(B) # inverse of B
```

### Solve linear systems
Matrices are related to systems of linear equations. 
Consider the two equations

$$
\begin{cases}
    x_1+3x_2  &= 7 \\
    2x_1+4x_2 &= 10
\end{cases}
$$
The system can be written in matrix form
$$\mathbf{A}\mathbf{x}=\mathbf{b}$$
where $\mathbf{A}$ is a matrix $\left[\begin{array}{rr} 1 & 3 \\ 2 & 4 \end{array}\right]$, $\mathbf{x}$ is a vector of unknown variables, $\left[\begin{array} {r}x_1\\ x_2  \end{array}\right]$,
and $\mathbf{b}$ is a vector of values of RHS, $\left[\begin{array} {r}7\\ 10  \end{array}\right]$.

Since $\mathbf{A}^{-1}\mathbf{A}=\mathbf{I}$ and $\mathbf{Ix}=\mathbf{x}$ we have
$$\mathbf{x} = \mathbf{A}^{-1}\mathbf{b} = \left[\begin{array}{rr} -2 & 1.5 \\ 1 & -0.5 \end{array}\right]\left[\begin{array} {r}7\\ 10  \end{array}\right] = \left[\begin{array} {r}1\\ 2  \end{array}\right]$$
The matrix $\mathbf{A}$ must be a unique, or full rank.

```{r}
A <- matrix(c(1,2,3,4),2,2)
b <- c(7,10)
x <- solve(A)%*%b; x
solve(A,b)
```

### Determinant
$2\times 2$:
$$\begin{vmatrix}a & b\\
c & d
\end{vmatrix} = a\cdot d - b\cdot c$$
$3\times 3$
$$\begin{vmatrix}
a & b & c\\
d & e & f \\
g & h & i
\end{vmatrix} = a \begin{vmatrix}e & f\\h & i\end{vmatrix} 
- b\begin{vmatrix}d & f\\g & i\end{vmatrix} 
+ c\begin{vmatrix}d & e\\g & h\end{vmatrix}$$

```{r}
det(A)
```

### Some other rule of Matrices
$$\begin{cases}

(A + B)^T &= A^T + B^T  \\
(AB)^T &= B^TA^T  \\
A(B + C) &= AB + AC \\ 
AB &= AC \nRightarrow B = C
\end{cases}
$$
In general $AB \neq BA$.
If $\alpha$ is a number then $\alpha AB = A(\alpha B)$.

### Eigenvalues and Eigenvectors 

```{r}
eigen(A)
```