---
title: DGLM (Poisson Distribution) - State-Space Model
author: Akihiko Mori
date: '2022-06-04'
slug: []
categories: []
tags:
  - Bayesian Modelling (R and Stan)
---
## Purpose
Bayesian estimation of the DGLM (Poisson distribution), a state space model, 
will be implemented using R and Stan. 
The implementation is done from data creation, 
so it can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. Stan
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(bayesplot)
library(gridExtra)

set.seed(2)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Time-Varying Coefficients Model|$\mu_t \sim Norm\left(\mu_{t-1},\sigma^2_w\right)$<br />$\gamma_t = Norm\left(\sum^{t-1}_{k=t-(K-1)}\gamma_t,\sigma^2_s\right)$<br />$\lambda_t=exp\left(\mu_t+\gamma_t\right)$<br />$y_t \sim Poiss\left(\gamma_t\right)$<br />|
| Simulation|$\mu_t \sim Norm\left(\mu_{t-1},0.3^2\right)$<br />$\gamma_t = Norm\left(\sum^{t-1}_{k=t-(K-1)}\gamma_t,0.1^2\right)$<br />$\lambda_t=exp\left(\mu_t+\gamma_t\right)$<br />$y_t \sim Poiss\left(\gamma_t\right)$<br />|

## Data Generation
```{r}
# Data Generation
time <- seq(as.POSIXct("2021/05/01"), as.POSIXct("2021/08/01"), "days")
y  <- c()
mu <- c()
gamma <- c()
mu[1] <- 0
gamma[1:6] <- c(-7,-5,2,3,6,9)/10 
T <- length(time)

for(t in 2:T){
  mu[t] <- rnorm(1, mu[t-1], 0.3)
}
for(t in 7:T){
  gamma[t] <- rnorm(1, -sum(gamma[(t-6):(t-1)]), 0.1)
}
lambda <- exp(mu + gamma)
for(t in 1:T){
  y[t] <- rpois(1, lambda[t])
}
data <- data.frame(time, y)
data %>% head()

## Data Visualization
plot <- data %>% 
  ggplot(aes(x=time)) +
  theme_classic(base_family = "") +
  theme(text = element_text(size = 10))+
  scale_x_datetime(date_labels = "%m/%d")

plot + 
  geom_point(aes(y=y))+
  geom_line(aes(y=y)) +
  labs(x="Time",y="Output",title="Output Trend")

## Parameter Visualization
plot_lambda_sim <- plot + 
  geom_point(aes(y=lambda))+
  geom_line(aes(y=lambda)) +
  labs(x="Time",y="λ",title="λ Trend")

plot_mu_sim <- plot + 
  geom_point(aes(y=mu))+
  geom_line(aes(y=mu)) +
  labs(x="Time",y="μ",title="μ Trend")

plot_gamma_sim <- plot + 
  geom_point(aes(y=gamma))+
  geom_line(aes(y=gamma)) +
  labs(x="Time",y="γ",title="γ Trend")

grid.arrange(plot_lambda_sim, plot_mu_sim, plot_gamma_sim)
```

## Stan
### Stan File
```{stan output.var="DGLMpois.stan"}
data {
  int T;
  int y[T];
}

parameters {
  vector[T] mu;
  vector[T] gamma;
  real<lower=0> sigma_w;
  real<lower=0> sigma_s;
  real<lower=0> sigma_v;
}

transformed parameters{
  vector[T] lambda;
  lambda = exp(mu+gamma);
}

model {
  mu[2:T] ~ normal(mu[1:(T-1)], sigma_w);
  for(t in 7:T){
    gamma[t] ~ normal(-sum(gamma[(t-6):(t-1)]), sigma_s);
  }
  y ~ poisson(lambda);
}

generated quantities{
  int y_pred[T];
  for(t in 1:T){
    y_pred[t] = poisson_rng(lambda[t]);
  }
}
```

### R for Stan
```{r}
data_list <- list(
  T = nrow(data),
  y = data$y
)

mcmc_result <- stan(
  file="DGLMpois.stan",
  data=data_list,
  seed=1,
  iter = 2000, warmup = 200, chains = 3, thin=1
)
```

## Analysis Results
### R code
```{r}
# Result
print(mcmc_result, pars=c("sigma_w", "sigma_s"), 
      probs = c(0.025, 0.5, 0.975))

## Convergence Check
mcmc_sample <- rstan::extract(mcmc_result, permuted=FALSE)
mcmc_combo(mcmc_sample, pars=c("sigma_w","sigma_s"))

## Parameter results
mcmc_sample <- rstan::extract(mcmc_result)
func <- function(x){
  return (quantile(x, c(0.025, 0.5, 0.975)))
}

## λ result
lambda <- apply(mcmc_sample[["lambda"]], 2, func)
plot_lambda <- plot + 
  geom_line(aes(y=lambda[2,]), col="blue") + 
  geom_ribbon(aes(ymin=lambda[1,],ymax=lambda[3,]), 
              alpha=0.5, fill="gray", col="blue") +
  labs(x="Time",y="λ",title="λ Trend")

## μ result
mu <- apply(mcmc_sample[["mu"]], 2, func)
plot_mu <- plot + 
  geom_line(aes(y=mu[2,]), col="blue") + 
  geom_ribbon(aes(ymin=mu[1,],ymax=mu[3,]), 
              alpha=0.5, fill="gray", col="blue")+
  labs(x="Time",y="μ",title="μ Trend")

## y result
gamma <- apply(mcmc_sample[["gamma"]], 2, func)
plot_gamma <- plot + 
  geom_line(aes(y=gamma[2,]), col="blue") + 
  geom_ribbon(aes(ymin=gamma[1,],ymax=gamma[3,]), 
              alpha=0.5, fill="gray", col="blue")+
  labs(x="Time",y="γ",title="γ Trend")

gridExtra::grid.arrange(plot_lambda, plot_mu, plot_gamma)

## Prediction Distribution
y_pred <- apply(mcmc_sample[["y_pred"]], 2, func)

plot + 
  geom_point(aes(y=y))+
  geom_line(aes(y=y_pred[2,]), col="blue") + 
  geom_ribbon(aes(ymin=y_pred[1,],ymax=y_pred[3,]), 
              alpha=0.5, fill="gray", col="blue") +
  labs(x="Time",y="Output",title="Output Trend")
```