---
title: Bayesian Estimation from Gaussian Distribution ($\lambda$)
author: Akihiko Mori
date: '2022-05-17'
slug: []
categories: []
tags: 
- Bayes Statistics
---

## Purpose
I will implement Bayesian estimation of the parameter $\lambda$ of a Gaussian (Normal) distribution using R. 
For the prior distribution, specify a Gamma distribution and look for the posterior distribution. 
Implemented from the creation of data, it can be copied and pasted.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data
5. Posterior Distribution

## Library
```{r}
library(ggplot2)
set.seed(1)
```

## True Distribution
```{r}
mu <- 170
lambda <- 1/100
x <- seq(100 ,200, 1)
prob <- dnorm(x, mu, 1/sqrt(lambda))

ggplot()+
  geom_line(aes(x=x, y=prob))+
  labs(x="x", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 24))+
  ggeasy::easy_center_title()
```


## Prior Distribution
```{r}
a_pre <- 2
b_pre <- 1
x <- seq(0, 0.1, 0.001)
lambda_prob_pre <- dgamma(x, a_pre, b_pre)

plot <- ggplot()+
  geom_line(aes(x=x, y=lambda_prob_pre))+
  labs(x="λ", y="prob", title="Gamma Distribution")+
  theme_classic(base_family = "")+
  theme(text = element_text(size = 20))
plot
```

## Data
```{r}
data <- rnorm(30, mu, 1/sqrt(lambda))
N <- length(data)
```

## Post Distribution
```{r}
a_pos <- N/2 + a_pre
b_pos <- sum((data-mu)^2)/2 + b_pre
x <- seq(0, 0.1, 0.001)
lambda_prob_pos <- dgamma(x, a_pos, b_pos)

plot + geom_line(aes(x=x, y=lambda_prob_pos), col="blue")
```
