---
title: Local-Linear Trend Model - State-Space Model
author: Akihiko Mori
date: '2022-06-04'
slug: []
categories: []
tags:
  - Bayesian Modelling (R and Stan)
---

## Purpose
Bayesian estimation of a local-linear trend model, a state space model, 
will be implemented using R and Stan. 
The code can be reproduced by copy and paste since it is implemented from data creation.

## TOC
1. Library
2. Model
3. Data Generation
4. Stan
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(bayesplot)
library(gridExtra)

set.seed(2)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
|Local Linear Trend Model|$\delta_t \sim Norm\left(\delta_{t-1},\sigma^2_d\right)$<br />$\mu_t \sim Norm\left(\mu_{t-1}+\delta_{t-1},\sigma^2_w\right)$<br />$y_t \sim Norm\left(\mu_t,\sigma^2_v\right)$<br />|
| Simulation|$\delta_t \sim Norm\left(\delta_{t-1},2^2\right)$<br />$\mu_t \sim Norm\left(\mu_{t-1}+\delta_{t-1},5^2\right)$<br />$y_t \sim Norm\left(\mu_t,3^2\right)$<br />|

## Data Generation
```{r}
# Data Generation
time <- seq(as.POSIXct("2021/05/01"), as.POSIXct("2021/08/01"), "days")
y <- c()
mu <- c()
delta <- c()
mu[1] <- rnorm(1, 20, 1) %>% round(1)
delta[1] <- rnorm(1, 5, 1) %>% round(1)
T <- length(time)

for(t in 2:T){
  delta[t] <- rnorm(1, delta[t-1], 2)
  mu[t] <- rnorm(1, mu[t-1]+delta[t-1], 5)
}

for(t in 1:t){
  y[t] <- rnorm(1, mu[t], 3) 
}
data <- data.frame(time, y)
data %>% head()

## Data Visualization
plot <- data %>% 
  ggplot(aes(x=time)) +
  theme_classic(base_family = "") +
  theme(text = element_text(size = 20))+
  scale_x_datetime(date_labels = "%m/%d")

plot + 
  geom_point(aes(y=y))+
  geom_line(aes(y=y)) +
  labs(x="Time",y="Output",title="Output Trend")

## Parameters Visualization
plot_mu_sim <- plot + 
  geom_point(aes(y=mu))+
  geom_line(aes(y=mu)) +
  labs(x="Time",y="μ",title="μ Trend")

plot_delta_sim <- plot + 
  geom_point(aes(y=delta))+
  geom_line(aes(y=delta)) +
  labs(x="Time",y="δ",title="δ Trend")

grid.arrange(plot_mu_sim, plot_delta_sim)
```

## Stan
### Stan File
```{stan output.var="localLinearTrendModel.stan"}
data {
  int T;
  vector[T] y;
}

parameters {
  vector[T] mu;
  vector[T] delta;
  real<lower=0> sigma_w;
  real<lower=0> sigma_v;
  real<lower=0> sigma_d;
}

model {
  delta[2:T] ~ normal(delta[1:(T-1)], sigma_d);
  mu[2:T] ~ normal(mu[1:(T-1)]+delta[1:(T-1)], sigma_w);
  y ~ normal(mu, sigma_v);
}

generated quantities{
  vector[T] y_pred;
  for(t in 1:T){
    y_pred[t] = normal_rng(mu[t], sigma_v);
  }
}
```

### R for Stan
```{r}
data_list <- list(
  T = nrow(data),
  y = data$y
)

mcmc_result <- stan(
  file="localLinearTrendModel.stan",
  data=data_list,
  seed=1,
  iter = 2000, warmup = 200, chains = 3, thin=1
)
```

## Analysis Results
### R code
```{r}
# Result
print(mcmc_result, pars=c("sigma_w", "sigma_v", "sigma_d"), 
      probs = c(0.025, 0.5, 0.975))


## Convergence Check
mcmc_sample <- rstan::extract(mcmc_result, permuted=FALSE)
mcmc_combo(mcmc_sample, pars=c("sigma_w","sigma_v","sigma_d"))

## Parameter results
mcmc_sample <- rstan::extract(mcmc_result)
func <- function(x){
  return (quantile(x, c(0.025, 0.5, 0.975)))
}
## μ result
mu <- apply(mcmc_sample[["mu"]], 2, func)
plot_mu <- plot + 
  geom_point(aes(y=y))+
  geom_line(aes(y=mu[2,]), col="blue") + 
  geom_ribbon(aes(ymin=mu[1,],ymax=mu[3,]), 
              alpha=0.5, fill="gray", col="blue") +
  labs(x="Time",y="μ",title="μ Trend")

## δ result
delta <- apply(mcmc_sample[["delta"]], 2, func)
plot_delta <- plot + 
  geom_line(aes(y=delta[2,]), col="blue") + 
  geom_ribbon(aes(ymin=delta[1,],ymax=delta[3,]), 
              alpha=0.5, fill="gray", col="blue") +
  labs(x="Time",y="δ",title="δ Trend")

grid.arrange(plot_mu, plot_delta)

# Prediction Distribution
y_pred <- apply(mcmc_sample[["y_pred"]], 2, func)

plot + 
  geom_point(aes(y=y))+
  geom_line(aes(y=y_pred[2,]), col="blue") + 
  geom_ribbon(aes(ymin=y_pred[1,],ymax=y_pred[3,]), 
              alpha=0.5, fill="gray", col="blue") +
  labs(x="Time",y="output",title="Prediction Distribution")
```