---
title: Bayeian Estimation  with Gaussian Distribution (μ)
author: Akihiko Mori
date: '2022-06-22'
slug: []
categories: []
tags:
  - Bayesian Statistics (R)
---

## Purpose
Bayesian estimation of the parameter μ of the Gaussian (Normal) distribution will be implemented using R. 
For the prior distribution, a Gaussian distribution will be specified, and the posterior distribution will be obtained. 
This will implement this from data generation, so it can be reproduced by copy and paste.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r message=FALSE}
library(ggplot2)
set.seed(1)
```

## True Distribution
```{r}
mu <- 170
lambda <- 1/100
x <- seq(100 ,200, 1)
prob <- dnorm(x, mu, 1/sqrt(lambda))

ggplot()+
  geom_line(aes(x=x, y=prob))+
  labs(x="x", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))
```


## Prior Distribution
```{r}
mu_pre <- 150
lambda_pre <- 1/10000
x <- seq(100 ,200, 1)
mu_prob_pre <- dnorm(x, mu_pre, 1/sqrt(lambda_pre))

plot <- ggplot()+
  geom_line(aes(x=x, y=mu_prob_pre))+
  labs(x="μ", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))
plot
```

## Data Generation
```{r}
data <- rnorm(30, mu, 1/sqrt(lambda))
N <- length(data)
```

## Posterior Distribution

```{r}
lambda_pos <- N*lambda + lambda_pre
mu_pos <- (lambda*sum(data) + lambda_pre*mu_pre)/lambda_pos
x <- seq(100 ,200, 1)
mu_prob_pos <- dnorm(x, mu_pos, 1/sqrt(lambda_pos))

plot + geom_line(aes(x=x, y=mu_prob_pos), col="blue")
```