---
title: Poisson Linear Model with BRMS
author: Akihiko Mori
date: '2022-05-29'
slug: []
categories: []
tags: 
- Bayesian Modelling (R and Stan)
---

## Purpose
Bayesian estimation of the GLM, a Poisson linear model, will be implemented using R and Stan. This code is implemented with the library of BRMS.
The data will be implemented from the data preparation and can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. BRMS
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(brms)
library(patchwork)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Normal Linear Model | $\mu_n=\beta_0+\beta_1x_{n1}+\cdots +\beta_{M}x_{nM}$ <br />$p(y_n | \beta,\sigma) \sim Normal(y_n|\mu_n,\sigma^2)$<br /> |
| This Model          | $\mu_n=\beta_0+\beta_1Age_n+\beta_2Qual_n$ <br />$p(Income_n | \beta,\sigma) \sim Normal(Income_n | \mu_n,\sigma^2)$<br />   |
| Simulation          |$\mu_n=300+10Age_n+200Qual_n$ <br />$Income_n \sim Normal(Income_n | \mu_n,100^2)$<br /> |

## Data Generation
```{r}
temp <- rnorm(100, 20,5) %>% round(1) 
holiday <- rbinom(100, 1, 2/7)
lambda <- exp(-2+0.2*temp+0.5*holiday)
sales <- rpois(100, lambda)
data　<- data.frame(temp, holiday, sales)
data %>% head()

plot <- ggplot() +
  geom_point(aes(x=data$temp, y=data$sales, color=factor(data$holiday))) + 
  theme_classic(base_family = "") +
  theme(text=element_text(size=20))+
  labs(x="Temperature", y="Sale counts", title="Scatter Plot") +
  scale_color_manual("Holiday",values=c("red","blue"))
plot
```

## BRMS
```{r}
mcmc_result <- brm(
  data = data,
  formula = sales ~ temp + holiday,
  family = poisson,
  seed = 1,
  iter = 1000, warmup = 200, chains = 4, thin=1
)
```


## Analysis Results
### R code
```{r}
# Result
print(mcmc_result)

## Convergence Check
theme_set(theme_classic(base_size = 10, base_family = ""))
plot(mcmc_result)

## mu result
plot(conditional_effects(mcmc_result, 
                         effects="temp:holiday"), 
     points=TRUE) %>% wrap_plots() +
  plot_annotation(title="λ's Estimation")

## Prediction distribution
plot(conditional_effects(mcmc_result, 
                         effects="temp:holiday", 
                         method="predict"),
     points=TRUE) %>% wrap_plots() +
  plot_annotation(title="prediction distribution")
```