---
title: Bayesian Estimation from Multivariate Normal Distribution ($\Lambda$)
author: Akihiko Mori
date: '2022-05-17'
slug: []
categories: []
tags: 
- Bayes Statistics
---

## Purpose
Bayesian estimation of the parameter $\lambda$ from multivariate Gaussian distribution (multivariate normal distribution) will be implemented using R. 
The prior distribution is assumed as Wishart distribution  and 
seek the posterior distribution. 
Implemented from the generation of data, the R code can be copied and pasted.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r}
library(ggplot2)
library(ellipse)
library(MCMCpack)
set.seed(1)
```

## True Distribution
```{r}
mu <- c(170, 60)
lambda <- solve(matrix(c(100, 80, 80, 100), ncol = 2))
samples <- mvrnorm(100, mu = mu, Sigma = solve(lambda))

ell <- ellipse(centre=mu, solve(lambda))
ggplot()+
  geom_point(aes(x=samples[,1], y=samples[,2]))+
  labs(x=expression(x[1]), y=expression(x[2]), title="Multivariate Normal Distribution")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 24))+
  lims(x=c(140,200), y=c(30,90))+
  geom_polygon(data=as.data.frame(ell), aes(x,y), fill=NA,colour="black")+
  ggeasy::easy_center_title()
```


## Prior Distribution
```{r}
n_pre <- 3
w_pre <- diag(2)*0.01
lambda_sample_pre <- rwish(n_pre, w_pre)

plot <- ggplot() +
  theme_classic(base_family = "")+
  theme(text = element_text(size = 20), legend.position=c(0.3,0.7)) +
  labs(x=expression(x[1]), y=expression(x[2]), title="Wishart distribution") 
for(i in 1:100){
  lambda_samples_pre <- rwish(n_pre, w_pre)
  ell <-  ellipse(centre=mu, solve(lambda_samples_pre))
  plot <- plot + geom_polygon(data=as.data.frame(ell),aes(x,y), fill=NA,colour="black")
}
plot
```

## Data Generation
```{r}
data <- mvrnorm(100, mu = mu, Sigma = solve(lambda))
N <- nrow(data)
```

## Post Distribution
```{r}
n_pos <- N + n_pre
S <- 0
for(i in 1:N){
  S <- S + (data[i,]-mu) %*% t(data[i,]-mu)
}
w_pos <- solve(S + solve(w_pre))

for(i in 1:100){
  lambda_samples_pos <- rwish(n_pos, w_pos)
  ell <-  ellipse(centre=mu, solve(lambda_samples_pos))
  plot <- plot + geom_polygon(data=as.data.frame(ell),aes(x,y), fill=NA, colour="blue")
}
plot
```
