---
title: Bayeian Estimation with Gaussian Distribution (μ,λ)
author: Akihiko Mori
date: '2022-06-22'
slug: []
categories: []
tags:
  - Bayesian Statistics (R)
---

## Purpose
Bayesian estimation of the parameter μ,λ of the Gaussian (Normal) distribution will be implemented using R. 
For the prior distribution, Gaussian and Gamma distributions will be specified, and the posterior distribution will be obtained. 
This will implement this from data generation, so it can be reproduced by copy and paste.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data Generation
5. Posterior Distribution

## Library
```{r message=FALSE}
library(ggplot2)
set.seed(1)
```

## True Distribution
```{r}
mu <- 170
lambda <- 1/100
x <- seq(100 ,200, 1)
prob <- dnorm(x, mu, 1/sqrt(lambda))

ggplot()+
  geom_line(aes(x=x, y=prob))+
  labs(x="x", y="prob", title="Gaussian Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))
```


## Prior Distribution
### Gamma Distribution
```{r}
a_pre <- 2
b_pre <- 1
x <- seq(0 ,0.1, 0.001)
lambda_prob_pre <- dgamma(x, a_pre, b_pre)
lambda_sample_pre <- rgamma(1, a_pre, b_pre)

plot_lambda <- ggplot()+
  geom_line(aes(x=x, y=lambda_prob_pre))+
  labs(x="λ", y="prob", title="Gamma Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3")+
  theme(text = element_text(size = 24))
plot_lambda
```

### Gaussian Distribution
```{r}
mu_pre <- 150
beta_pre <- 0.0001
x <- seq(100 ,200, 1)
mu_prob_pre <- dnorm(x, mu_pre, 1/sqrt(beta_pre*lambda_sample_pre))

plot_mu <- ggplot()+
  geom_line(aes(x=x, y=mu_prob_pre))+
  labs(x="μ", y="prob", title="Gaussian Dtribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))
plot_mu
```


## Data Generation
```{r}
data <- rnorm(30, mu, 1/sqrt(lambda))
N <- length(data)
```

## Posterior Distribution
### Parameter Update
```{r}
beta_pos <- N + beta_pre
mu_pos <- (sum(data)+beta_pre*mu_pre)/beta_pos
a_pos <- N/2 + a_pre
b_pos <- (sum(data^2)+beta_pre*mu_pre^2-beta_pos*mu_pos^2)/2 + b_pre
lambda_sample_pos <- rgamma(1, a_pos, b_pos)
```

### Gamma Distribution
```{r}
x <- seq(0 ,0.1, 0.001)
lambda_prob_pos <- dgamma(x, a_pos, b_pos)

plot_lambda + geom_line(aes(x=x, y=lambda_prob_pos), col="blue")
```

### Gaussian Distribution
```{r}
x <- seq(100 ,200, 1)
mu_prob_pos <- dnorm(x, mu_pos, 1/sqrt(beta_pos*lambda_sample_pos))

plot_mu + geom_line(aes(x=x, y=mu_prob_pos), col="blue")
```