---
title: Logistic Linear Model with BRMS
author: Akihiko Mori
date: '2022-06-01'
slug: []
categories: []
tags:
  - Bayesian Modelling (R and Stan)
---
## Purpose
Bayesian estimation of the GLM, a Logistic linear model, will be implemented using R and Stan. This code is implemented with the library of BRMS.
The data will be implemented from the data preparation and can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. BRMS
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(brms)
library(patchwork)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Poisson Linear Model | $p_n=sigmoid\left(\beta_0+\beta_1x_{n1}+\cdots +\beta_{M}x_{nM}\right)$ <br />$p\left(y_n | \beta\right) \sim Bern\left(y_n|p_n\right)$<br /> |
| This Model          | $p_n=sigmoid\left(\beta_0+\beta_1Age_n+\beta_2Qual_n\right)$ <br />$p\left(Pass_n | \beta\right) \sim Bern\left(Pass_n | p_n\right)$<br />   |
| Simulation          |$p_n=sigmoid\left(10-0.4Age_n+7Qual_n\right)$ <br />$Pass_n \sim Bern\left(Pass_n | p_n\right)$<br /> |

## Data Generation
```{r}
qual <- rbinom(100, 1, 0.3)
age <- rnorm(100, 50, 20) %>% round()
p <- 1/(1+exp(-(10-0.4*age + 7*qual)))
pass <- rbinom(100, 1, p)
data <- data.frame(qual, age, pass) %>% filter(age>18, age<65)
data %>% head()

plot <- ggplot() +
  geom_point(aes(x=data$age, y=data$pass, color=factor(data$qual))) + 
  theme_classic(base_family = "") +
  theme(text=element_text(size=20))+
  labs(x="Age", y="Pass", title="Scatter plot") +
  scale_color_manual("Qualification",values=c("red","blue"))+
  ggeasy::easy_center_title()
plot
```

## BRMS
```{r}
mcmc_result <- brm(
  data = data,
  formula = pass ~ age + qual,
  family = bernoulli(),
  seed = 1,
  iter = 1000, warmup = 200, chains = 4, thin=1
)
```


## Analysis Results
### R code
```{r}
# Result
print(mcmc_result)

## Convergence Check
theme_set(theme_classic(base_size = 10, base_family = "HiraKakuProN-W3"))
plot(mcmc_result)

## mu result
plot(conditional_effects(mcmc_result, 
                         effects="age:qual"), 
     points=TRUE) %>% wrap_plots() +
  plot_annotation(title="p's Estimation")

## Prediction distribution
plot(conditional_effects(mcmc_result, 
                         effects="age:qual", 
                         method="predict"),
     points=TRUE) %>% wrap_plots() +
  plot_annotation(title="prediction distribution")
```