---
title: Local Level Model - State-Space Model
author: Akihiko Mori
date: '2022-06-03'
slug: []
categories: []
tags:
  - Bayesian Modelling (R and Stan)
---
## Purpose
Bayesian estimation of a local level model, a state space model, 
will be implemented using R and Stan. 
The code can be reproduced by copy and paste since it is implemented from data creation.

## TOC
1. Library
2. Model
3. Data Generation
4. Stan
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(bayesplot)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Local Level Model| $\mu_t \sim Norm\left(\mu_{t-1},\sigma^2_w\right)$ <br /> $y_t \sim Norm\left(\mu_t,\sigma^2_v\right)$<br />|
| Simulation| $\mu_t \sim Norm\left(\mu_{t-1},5^2_w\right)$ <br /> $y_t \sim Norm\left(\mu_t,3^2_v\right)$<br />|

## Data Generation
```{r}
time  <- seq(as.POSIXct("2021/05/01"), as.POSIXct("2021/08/01"), "days")
y     <- c()
mu    <- c()
mu[1] <- rnorm(1, 200, 10) %>% round(1)
T <- length(time)
for(t in 2:T){
  mu[t] <- rnorm(1, mu[t-1], 5)
}
for(t in 1:T){
  y[t] <- rnorm(1, mu[t], 3) 
}
data <- data.frame(time, y)
data %>% head()

plot <- ggplot() +
  geom_point(aes(x=data$time, y=data$y))+
  geom_line(aes(x=data$time, y=data$y))+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 30))+
  labs(x="Time",y="Y",title="Trend of Y") +
  scale_x_datetime(date_labels = "%m/%d")
plot
```

## Stan
### Stan File
```{stan output.var="localLevelModel.stan"}
data {
  int T;
  vector[T] y;
  
  int T_pred;
}

parameters {
  vector[T] mu;
  real<lower=0> sigma_w;
  real<lower=0> sigma_v;
}

model {
  mu[2:T] ~ normal(mu[1:(T-1)], sigma_w);
  y ~ normal(mu, sigma_v);
}

generated quantities{
  vector[T+T_pred] mu_pred;
  vector[T+T_pred] y_pred;
  
  mu_pred[1:T] = mu;
  for(t in 1:T_pred){
    mu_pred[T+t] = normal_rng(mu_pred[T+t-1], sigma_w);
  }
  for(t in 1:(T+T_pred)){
    y_pred[t] = normal_rng(mu_pred[t], sigma_v);
  }
}
```

### R for Stan
```{r}
data_list <- list(
  T = nrow(data), K = 4, y = data$y,
  T_pred = 10
)

mcmc_result <- stan(
  file="localLevelModel.stan",
  data=data_list,
  seed=10,
  iter = 2000, warmup = 200, chains = 3, thin=1
)
```

## Analysis Results
### R code
```{r}
# Result
print(mcmc_result, pars=c("sigma_w", "sigma_v"), probs = c(0.025, 0.5, 0.975))

## Convergence Check
mcmc_sample <- rstan::extract(mcmc_result, permuted=FALSE)
mcmc_combo(mcmc_sample, pars=c("sigma_w","sigma_v"))

## μ result
mcmc_sample <- rstan::extract(mcmc_result)
func <- function(x){
  return (quantile(x, c(0.025, 0.5, 0.975)))
}
mu_pred <- apply(mcmc_sample[["mu_pred"]], 2, func)
Time_pred <- seq(as.POSIXct("2021/05/01"), as.POSIXct("2021/08/11"), "days")

plot + 
  labs(title="μ Estimation") +
  geom_line(aes(x=Time_pred, y=mu_pred[2,]), col="blue") + 
  geom_ribbon(aes(x=Time_pred, ymin=mu_pred[1,], ymax=mu_pred[3,]), 
              alpha=0.5, fill="gray", col="blue")

# Prediction Distribution
y_pred <- apply(mcmc_sample[["y_pred"]], 2, func)

plot +
  labs(title="Prediction Distribution") +
  geom_line(aes(x=Time_pred, y=y_pred[2,]), col="blue") + 
  geom_ribbon(aes(x=Time_pred, ymin=y_pred[1,], ymax=y_pred[3,]), 
              alpha=0.5, fill="gray", col="blue")
```