---
title: Normal Linear Model
author: Akihiko Mori
date: '2022-05-18'
slug: []
categories: []
tags: 
- Bayesian Modelling (R and Stan)
---
## Purpose
Bayesian estimation of the GLM, a normal linear model, will be implemented using R and Stan.
The data will be implemented from the data preparation and can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. Stan
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(bayesplot)
library(ggpubr)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Normal Linear Model | $\mu_n=\beta_0+\beta_1x_{n1}+\cdots +\beta_{M}x_{nM}$ <br />$p(y_n | \mu_n,\sigma) \sim Normal(y_n|\mu_n,\sigma^2)$<br /> |
| This Model          | $\mu_n=beta_0+\beta_1Age_n+\beta_2Qual_n$ <br />$p(Income_n | \mu,\sigma) \sim Normal(Income_n | \mu_n,\sigma^2)$<br />   |
| Simulation          |$\mu_n=300+10Age_n+200Qual_n$ <br />$p(Income_n|\mu,\sigma) \sim Normal(Income_n | \mu_n,100^2)$<br /> |

## Data Generation
```{r}
qual <- rbinom(100, 1, 0.3)
age <- rnorm(100, 50, 20) %>% round()
income <- rnorm(100, 300 + 10*age + 200*qual, 100) %>% round()
data <- data.frame(qual, age, income) %>% filter(age>18, age<65)
data %>% head()

plot <- ggplot() +
  geom_point(aes(x=data$age, y=data$income, color=factor(data$qual))) + 
  theme_classic(base_family = "") +
  theme(text=element_text(size=20))+
  labs(x="Age", y="Icome", title="Model") +
  scale_color_manual("Qualification",values=c("red","blue"))+
  ggeasy::easy_center_title()
plot
```

## Stan
### Stan File
```{stan output.var="NormalLinearModel.stan"}
data {
  int N;
  vector[N] y;
  vector[N] x1;
  vector[N] x2;
  
  int N_hat;
  vector[N_hat] x_hat;
}

parameters {
  vector[3] b;
  real<lower=0> sigma;
}

transformed parameters{
  vector[N] mu;
  mu = b[1] + b[2]*x1 + b[3]*x2;
}

model{
  y ~ normal(mu, sigma);
}

generated quantities{
  vector[N_hat] mu_hat1;
  vector[N_hat] mu_hat0;
  vector[N_hat] y_hat1;
  vector[N_hat] y_hat0;
  
  for(n in 1:N_hat){
    mu_hat1[n] = b[1] + b[2]*x_hat[n] + b[3];
    mu_hat0[n] = b[1] + b[2]*x_hat[n];
    y_hat1[n] = normal_rng(mu_hat1[n], sigma);
    y_hat0[n] = normal_rng(mu_hat0[n], sigma);
  }
}
```


### R for Stan
```{r}
x_hat <- seq(min(data$age), max(data$age))
data_list <- list(
  N = nrow(data), y = data$income,
  x1 = data$age, x2 = data$qual,
  
  N_hat = length(x_hat), x_hat = x_hat
)

mcmc_result <- stan(
  file="NormalLinearModel.stan",
  data=data_list,
  seed=1,
  iter = 1000, warmup = 200, chains = 4, thin=1
)
```

## Analysis Results
### R code
```{r}
# Result
print(mcmc_result, probs = c(0.025, 0.5, 0.975), pars=c("b","sigma"))

## Convergence Check
mcmc_sample <- rstan::extract(mcmc_result, permuted=FALSE)
mcmc_combo(mcmc_sample, pars=c("b[1]","b[2]","b[3]","sigma"))

## mu result
mcmc_sample <- rstan::extract(mcmc_result)
func <- function(x){
  return (quantile(x, c(0.025, 0.5, 0.975)))
}
mu_hat1 <- apply(mcmc_sample[["mu_hat1"]], 2, func)
mu_hat0 <- apply(mcmc_sample[["mu_hat0"]], 2, func)

plot_mu <- plot + 
  labs(title="Estimation of μ") +
  geom_line(aes(x=x_hat, y=mu_hat1[2,]), col="blue") + 
  geom_line(aes(x=x_hat, y=mu_hat0[2,]), col="red") +
  geom_ribbon(aes(x=x_hat, ymin=mu_hat1[1,],ymax=mu_hat1[3,]), alpha=0.5, fill="gray", col="blue")+
  geom_ribbon(aes(x=x_hat, ymin=mu_hat0[1,],ymax=mu_hat0[3,]), alpha=0.5, fill="gray", col="red")
plot_mu

## Prediction distribution
y_hat1 <- apply(mcmc_sample[["y_hat1"]], 2, func)
y_hat0 <- apply(mcmc_sample[["y_hat0"]], 2, func)

plot_y <- plot +
  labs(title="Prediction Distribution") +
  geom_line(aes(x=x_hat, y=y_hat1[2,]), col="blue") + 
  geom_line(aes(x=x_hat, y=y_hat0[2,]), col="red") +
  geom_ribbon(aes(x=x_hat, ymin=y_hat1[1,],ymax=y_hat1[3,]), alpha=0.5, fill="gray", col="blue")+
  geom_ribbon(aes(x=x_hat, ymin=y_hat0[1,],ymax=y_hat0[3,]), alpha=0.5, fill="gray", col="red")
plot_y
```

