---
title: Generalized Linear Mixed Model (Random Intercept 2)
author: Akihiko Mori
date: '2022-06-02'
slug: []
categories: []
tags:
  - Bayesian Modelling (R and Stan)
---
## Purpose
Bayesian estimation of the GLM, a Generalized Linear Mixed Model - Random Intercept - each variable, will be implemented using R and Stan.
The data will be implemented from the data preparation and can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. Stan
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(bayesplot)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Generalized Linear Mixed Model (Random Intercept) | $\lambda_n=exp\left(\beta_0+\beta_1x_{n1}+\cdots +\beta_{M}x_{nM}+\textbf{r}_k\right)$ <br />$p\left(y_n | \beta\right) \sim Poiss\left(y_n|\lambda_n\right)$<br /> |
| This Model          | $\lambda_n=exp\left(\beta_0+\beta_1Temp_n+\beta_2Holiday_n+\textbf{r}_k\right)$ <br />$p\left(y_n | \beta\right) \sim Poiss\left(y_n | \lambda_n\right)$<br />   |
| Simulation          |$\textbf{r}_k \sim Norm\left(0,0.6\right)$ <br /> $\lambda_n=exp\left(-2+0.2Temp_n+0.5Holiday_n+\textbf{r}_k\right)$<br /> $Sales_n \sim Poiss\left(Sales_n | \lambda_n\right)$<br /> |

## Data Generation
```{r}
temp <- rnorm(100, 20,5) %>% round(1) 
holiday <- rbinom(100, 1, 2/7)
shop <- runif(100, 1, 4) %>% round()
shopRandEffect <- rnorm(4, 0, 1)
data　<- data.frame(temp, holiday, shop) %>% 
  mutate(r=ifelse(shop==1, 
                  shopRandEffect[1], 
                  ifelse(shop==2, 
                         shopRandEffect[2],
                         ifelse(shop==3,
                                shopRandEffect[3],
                                shopRandEffect[4] ))))%>%
  mutate(lambda=exp(-2+0.2*temp+0.5*holiday+r))
data$sales <- rpois(100, data$lambda)
data %>% select(temp, holiday, shop, sales) %>% head()

plot <- ggplot() +
  geom_point(data=data,aes(x=temp, y=sales, color=factor(holiday))) + 
  theme_classic(base_family = "") +
  theme(text=element_text(size=20))+
  labs(x="Temperature", y="Sale Counts", title="Scatter Plot") +
  scale_color_manual("Qualification",values=c("red","blue")) + 
  facet_wrap(.~ shop)
plot
```

## Stan
### Stan File
```{stan output.var="GLM2int.stan"}
data {
  int N;
  int y[N];
  vector[N] x1;
  vector[N] x2;
  
  int K;
  int<lower=1, upper=K> x3[N];
}

parameters {
  vector[3] b;
  
  vector[K] r;
  real<lower=0> sigma;
}

transformed parameters{
  vector[N] lambda;
  lambda = exp(b[1] + b[2]*x1 + b[3]*x2 + r[x3]);
}

model{
  r ~ normal(0, sigma);
  y ~ poisson(lambda);
}
```


### R for Stan
```{r}
data_list <- list(
  N = nrow(data), K = 4, y = data$sales,
  x1 = data$temp, x2 = data$holiday, x3 = data$shop
)

mcmc_result <- stan(
  file="GLM2int.stan",
  data=data_list,
  seed=1,
  iter = 2000, warmup = 200, chains = 4, thin=1
)
```

## Analysis Results
### R code
```{r}
# Result
print(mcmc_result, probs = c(0.025, 0.5, 0.975), pars=c("b","sigma", "r"))

## Convergence Check
mcmc_sample <- rstan::extract(mcmc_result, permuted=FALSE)
mcmc_combo(mcmc_sample, pars=c("b[1]","b[2]","b[3]","sigma"))
```