---
title: Bayesian Estimation from Categorical Distribution
author: Akihiko Mori
date: '2022-05-16'
slug: []
categories: []
tags: 
- Bayes Statistics
---

## Purpose
I will implement Bayesian estimation of the parameter p of a Categorical distribution using R. 
For the prior distribution, specify a Direct distribution and look for the posterior distribution. 
Implemented from the creation of data, it can be copied and pasted.

## TOC
1. Library
2. True Distribution
3. Prior Distribution
4. Data
5. Post Distribution

## Library
```{r}
library(ggplot2)
library(MCMCpack) # rdirichlet
set.seed(1)
```

## True Distribution
```{r}
x <- c(1, 2, 3)
p <- c(0.1, 0.6, 0.3)
prob <- p

ggplot() + 
  geom_bar(aes(x, prob), stat ="identity") +
  labs(x="x", y="prob", title="Categorical Distribution")+
  theme_classic(base_family = "") +
  theme(text = element_text(size = 20))+
  ggeasy::easy_center_title()
```


## Prior Distribution
```{r}
alpha_pre <- c(1, 1, 1)
p_samples_pre <- rdirichlet(100, alpha_pre)

plot <- ggplot()+
  geom_point(aes(x=p_samples_pre[,1], y=p_samples_pre[,2]))+
  labs(x="1", y="2", title="Direct Distribution")+
  theme_classic(base_family = "HiraKakuPro-W3") +
  theme(text = element_text(size = 24))+
  lims(x=0:1, y=0:1)+
  ggeasy::easy_center_title()
plot
```

## Data
```{r}
data <- which(rmultinom(300, 1, c(0.1, 0.6, 0.3))==1, arr.ind = TRUE)[,1]
N <- length(data)
```

## Post Distribution
```{r}
alpha_pos <- as.vector(table(data)) + alpha_pre
p_samples_pos <- rdirichlet(100, alpha_pos)

plot + geom_point(aes(x=p_samples_pos[,1], y=p_samples_pos[,2]), col="blue")
```

