---
title: Poisson Linear model
author: Akihiko Mori
date: '2022-05-29'
slug: []
categories: []
tags: 
- Bayesian Modelling (R and Stan)
---

## Purpose
Bayesian estimation of the GLM, a Poisson linear model, will be implemented using R and Stan.
The data will be implemented from the data preparation and can be reproduced by copy and paste.

## TOC
1. Library
2. Model
3. Data Generation
4. Stan
5. Analysis

## Library
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(rstan)
library(bayesplot)

set.seed(1)
rstan_options(auto_write=TRUE)
options(mc.cores=parallel::detectCores())
```

## Model
| Model | Model Equation |
|-------|----------------------------------|
| Normal Linear Model | $\lambda_n=exp(\beta_0+\beta_1x_{n1}+\cdots +\beta_{M}x_{nM})$ <br />$p(y_n | \beta,\sigma) \sim Poiss(y_n|\lambda_n)$<br /> |
| This Model          | $\lambda_n=\beta_0+\beta_1Temp_n+\beta_2Holiday_n$ <br />$p(Sales_n | \beta) \sim Poiss(Sales_n | \lambda_n)$<br />   |
| Simulation          |$\lambda_n=-2+0.2Temp_n+0.5Holiday_n$ <br />$Sales_n \sim Poiss(Sales_n | \lambda_n)$<br /> |

## Data Generation
```{r}
temp <- rnorm(100, 20,5) %>% round(1) 
holiday <- rbinom(100, 1, 2/7)
lambda <- exp(-2+0.2*temp+0.5*holiday)
sales <- rpois(100, lambda)
data　<- data.frame(temp, holiday, sales)
data %>% head()

plot <- ggplot() +
  geom_point(aes(x=data$temp, y=data$sales, color=factor(data$holiday))) + 
  theme_classic(base_family = "") +
  theme(text=element_text(size=20))+
  labs(x="Temperature", y="Sale counts", title="Scatter Plot") +
  scale_color_manual("Holiday",values=c("red","blue"))
plot
```

## Stan
### Stan File
```{stan output.var="PoissonLinearModel.stan"}
data {
  int N;
  int y[N];
  vector[N] x1;
  vector[N] x2;
  
  int N_hat;
  vector[N_hat] x_hat;
}

parameters {
  vector[3] b;
}

transformed parameters{
  vector[N] lambda;
  lambda = exp(b[1] + b[2]*x1 + b[3]*x2);
}

model{
  y ~ poisson(lambda);
}

generated quantities{
  vector[N_hat] lambda_hat1;
  vector[N_hat] lambda_hat0;
  int y_hat1[N_hat];
  int y_hat0[N_hat];
  
  for(n in 1:N_hat){
    lambda_hat1[n] = exp(b[1] + b[2]*x_hat[n] + b[3]);
    lambda_hat0[n] = exp(b[1] + b[2]*x_hat[n]);
    y_hat1[n] = poisson_rng(lambda_hat1[n]);
    y_hat0[n] = poisson_rng(lambda_hat0[n]);
  }
}
```


### R for Stan
```{r}
x_hat <- seq(min(data$temp), max(data$temp))
data_list <- list(
  N = nrow(data), y = data$sales,
  x1 = data$temp, x2 = data$holiday,
  
  N_hat = length(x_hat), x_hat = x_hat
)

mcmc_result <- stan(
  file="PoissonLinearModel.stan",
  data=data_list,
  seed=10,
  iter = 2000, warmup = 200, chains = 3, thin=1
)
```

## Analysis Results
### R code
```{r}
# Result
print(mcmc_result, probs = c(0.025, 0.5, 0.975), pars="b")

## Convergence Check
mcmc_sample <- rstan::extract(mcmc_result, permuted=FALSE)
mcmc_combo(mcmc_sample, pars=c("b[1]","b[2]","b[3]"))

## lambda result
mcmc_sample <- rstan::extract(mcmc_result)
func <- function(x){
  return (quantile(x, c(0.025, 0.5, 0.975)))
}
lambda_hat1 <- apply(mcmc_sample[["lambda_hat1"]], 2, func)
lambda_hat0 <- apply(mcmc_sample[["lambda_hat0"]], 2, func)

plot_lambda<- plot + 
  labs(title="λ Estimation") +
  geom_line(aes(x=x_hat, y=lambda_hat1[2,]), col="blue") + 
  geom_line(aes(x=x_hat, y=lambda_hat0[2,]), col="red") +
  geom_ribbon(aes(x=x_hat, ymin=lambda_hat1[1,],ymax=lambda_hat1[3,]), alpha=0.5, fill="gray", col="blue")+
  geom_ribbon(aes(x=x_hat, ymin=lambda_hat0[1,],ymax=lambda_hat0[3,]), alpha=0.5, fill="gray", col="red")
plot_lambda

## Prediction distribution
y_hat1 <- apply(mcmc_sample[["y_hat1"]], 2, func)
y_hat0 <- apply(mcmc_sample[["y_hat0"]], 2, func)

plot_y <- plot + 
  labs(title="Prediction Distribution") +
  geom_line(aes(x=x_hat, y=y_hat1[2,]), col="blue") + 
  geom_line(aes(x=x_hat, y=y_hat0[2,]), col="red") +
  geom_ribbon(aes(x=x_hat, ymin=y_hat1[1,],ymax=y_hat1[3,]), alpha=0.5, fill="gray", col="blue")+
  geom_ribbon(aes(x=x_hat, ymin=y_hat0[1,],ymax=y_hat0[3,]), alpha=0.5, fill="gray", col="red")
plot_y
```